Maximum Entropy Spectral Analysis: a case study

### 2021-03-24 paper/mesa.tex
Added plots

\section{Application: Temperature Time Series} \label{sec:temp_timeseries}
We apply our code to the study of the temperature time series. This will allow to test the code performance on a different time series, of great importance in physics. We will also assess the quality of the forecasting predictions: this task has a great impact also on practical applications.
We downloaded the historical reanalysis data from the Copernicus project (or whathever it is) \cite{}.
Data refers to the temperature measure at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy. The temperature is measured once per hour for almost 31 years from 31st December 1989 to 30th November 2020.

@@

Looking at the low frequencies, it is striking to note that FPE does not capture the peak $f_{yr} = \SI{1}{yr^{-1}}$ located at the frequency of a year. On the other hand, the peak is captured well by the CAT loss function.
This can be understood by looking at fig.~\ref{fig:p_vs_ptrue}. It is shown that CAT systematically select very long filters (i.e. large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often understimating badly the actual value.
In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
Clearly, the autoregressive process selected by FPE cannot produce a meaningful prediction on the timescale of a few months and for this reason it is unable to capture the peak at $f_{yr}$. In other words, the model is too simple to capture \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour.
The ``actual" length of the autoregressive process should be closer to the choice of CAT.

@@

>>>>

\section{Applications} \label{sec:applications}
\subsection{Temperature Time Series} \label{sec:temp_timeseries}
We apply our code to the study of the temperature time series. This will allow to test the code performance on a different time series, of great importance in physics. We will also assess the quality of the forecasting predictions: this task has a great impact also on practical applications.
We downloaded the historical reanalysis data from the Copernicus project (or whathever it is) \cite{}.
Data refers to the temperature measure at coordinates $\textrm{N}\;\ang{45;5;} \;\; \textrm{E}\;\ang{9;1;}}$, corresponding about to the city of Milan, Italy. The temperature is measured once per hour for almost 31 years from 31st December 1989 to 30th November 2020.

@@

Looking at the low frequencies, it is striking to note that FPE does not capture the peak $f_{yr} = \SI{1}{yr^{-1}}$ located at the frequency of a year. On the other hand, the peak is captured well by the CAT loss function.
In analogy as what observed for the peak at $f_D$, in the CAT spectru we observe a higher order harmonic at a frequency $2 \cdot f_{yr}$. As in the previous case, this is required to better model the temperature variation in the year.

The failure of FPE in catching the low frequency trend can be understood by looking at fig.~\ref{fig:p_vs_ptrue}. It is shown that CAT systematically select very long filters (i.e. large $p_{CAT}$ values for the autoregressive orders), whereas FPE tends to select smaller values for the order $p_{FPE}$ of the autoregressive process, often understimating badly the actual value.
In this example, we have $p_{CAT} \sim  36500 \sim  \SI{4}{yr}$, whereas $p_{FPE} \sim  1300 \sim  \SI{2}{months}$.
Clearly, the autoregressive process selected by FPE cannot produce a meaningful prediction on the timescale of a few months and for this reason it is unable to capture the peak at $f_{yr}$. In other words, the model is too simple to capture \textit{both} the behaviour at high and low frequencies (separated by approximately 3 orders of magnitude).
On the other hand, the model chosen by CAT is much more complex and is able to model also the high frequency behaviour.
The ``actual" length of the autoregressive process should be closer to the choice of CAT.

@@

\subsection{Seismic waveforms} \label{sec:seismic_timeseries}
\blindtext
\begin{figure}
	\caption{Do a caption}
	\label{fig:seismic_spectrum}
	\includegraphics{Images/seismic_plots/seismic_spectrum.pdf}
\end{figure}

\begin{figure}
	\caption{Do a caption}
	\label{fig:seismic_forecast}
	\includegraphics{Images/seismic_plots/forecast_accuracy.pdf}
\end{figure}

\subsection{Forecasting the LIGO strain} \label{sec:LIGO_forecasting}
\blindtext
\begin{figure}
	\caption{Do a caption}
	\label{fig:LIGO_forecast}
	\includegraphics{Images/forecast_LIGO/forecast_accuracy.pdf}
\end{figure}
